Bow shocks, bow waves, and dust waves. IV. Bow shape statistics

### 2019-04-18 papers/Papers1234/obs-bowshocks.tex
Add Luis A to the author list for paper 3
Move the oscillating bows section to the observational paper
Finish intro to paper II
Start work on Paper IV - change Pi-Lambda graphs to log scale
Two more paras of introduction for Paper IV
Lots of reorganizing of Paper IV

% The following is needed to fix the margins if using Letter-size paper
% REMOVE if your LaTeX uses A4 paper by default
%\addtolength\topmargin{-1.8cm}

% Standard LaTeX packages
% \usepackage[varg]{txfonts}
\usepackage[varg]{newtxmath}
\usepackage{newtxtext}
\usepackage{graphicx}
\usepackage{microtype}
\usepackage{xcolor}
\usepackage{fixltx2e}
\usepackage{booktabs}
\usepackage{siunitx}
\usepackage{color}
\usepackage{enumerate}
\usepackage{pdflscape}
\usepackage{rotating}
\usepackage{hyperref}

% Define a fermata symbol for use inline with text
% 04 Jul 2017 WJH
% Based on info in Secs 18.4.1 and 24.1 of musixdoc.pdf
\usepackage{musixtex}
\makeatletter
\newcommand\textfermata{%
  {\let\extractline\relax
    \setlines10\smallmusicsize \nobarnumbers \nostartrule
    \staffbotmarg0pt \setclefsymbol1\empty \global\clef@skip0pt
    \raisebox{0ex}[0ex][0ex]{
      \startextract\addspace{-\afterruleskip}%
      \notes\fermataup{-2}\en
      \zendextract}}}
\makeatother

@@

% Definitions needed in abstract
\newcommand\hii{\ion{H}{ii}}

\title[Stellar bow shocks]{Weather vanes and runaways: Bow
  shocks across the HR diagram}

\newcommand\AddressCRyA{Instituto de Radioastronom\'{\i}a y Astrof\'{\i}sica,
  Universidad Nacional Aut\'onoma de M\'exico, Apartado Postal 3-72,
  58090 Morelia, Michoac\'an, M\'exico}
\author[Henney \& Tarango Yong]{
  William J. Henney \& Jorge A. Tarango Yong \\
  \AddressCRyA
}
\begin{document}
\maketitle
\begin{abstract}
  Stellar bow shocks are the result of the supersonic interaction
  between a stellar wind and its environment.  Some of these are
  "runaways": high-velocity stars that have been ejected from a star
  cluster.  Others are "weather vanes", where it is the local
  interstellar medium itself that is moving, perhaps as the result of
  a champagne flow of ionized gas from a nearby \hii{} region.
  We recently proposed a new two-dimensional classification scheme for
  the shape of such bow shocks, which we here apply to three very
  different observational datasets: (1)~mid-infrared arcs
  around hot main-sequence stars (\(N = 227\)); (2)~far-infrared arcs
  around luminous cool stars (\(N = 7\)); and (3)~emission-line arcs
  around proplyds and other young stars in the Orion Nebula
  (\(N = 18\)).  We find significant differences between the three
  datasets: cool star bow shocks have markedly more closed wings than
  hot star bow shocks, while differences in the shape of the apex
  region are only marginally significant.  The Orion Nebula arcs, on
  the other hand, have both significantly more open wings and
  significantly flatter apices than the hot star bow shocks.  We
  discuss the implications of these differences for understanding the
  physics of the bow shock interaction in the three classes of
  sources.
\end{abstract}

\input{sec-obs-intro}
\clearpage
\input{sec-quadrics-observations}
\input{sec-obs-conclusions}
\clearpage
\bibliographystyle{mnras}
%% All references should be put in the BibTeX file bowshocks-biblio.bib
\bibliography{bowshocks-biblio}
\appendix
\input{app-p-values}

>>>>

% Standard LaTeX packages
\usepackage{graphicx}
\usepackage{microtype}
\usepackage{xcolor}
\usepackage{fixltx2e}
\usepackage{booktabs}
\usepackage{siunitx}
\usepackage{color}
\usepackage{enumerate}
\usepackage{pdflscape}
\usepackage{rotating}
\usepackage{hyperref}

\usepackage[T1]{fontenc} 
\usepackage[utf8]{inputenc}

% Fonts 
\usepackage{newtxtext}
% Note: newtxmath must come AFTER newtxtext
\usepackage[varvw,smallerops]{newtxmath}

\usepackage{chemgreek}
\activatechemgreekmapping{newtx}

% Define a fermata symbol for use inline with text
% 04 Jul 2017 WJH
% Based on info in Secs 18.4.1 and 24.1 of musixdoc.pdf
\let\ORIGna\na
\usepackage{musixtex}
\makeatletter
\newcommand\textfermata{%
  {\let\extractline\relax
    \setlines10\smallmusicsize \nobarnumbers \nostartrule
    \staffbotmarg0pt \setclefsymbol1\empty \global\clef@skip0pt
    \raisebox{0ex}[0ex][0ex]{
      \startextract\addspace{-\afterruleskip}%
      \notes\fermataup{-2}\en
      \zendextract}}}
\makeatother
\let\na\ORIGna

@@

\defcitealias{Tarango-Yong:2018a}{Paper~0}
\newcommand\PaperI{\citetalias{Tarango-Yong:2018a}}

% \title[Stellar bow shocks]{Weather vanes and runaways: Bow
%   shocks across the HR diagram}
\title[Bow shocks, bow waves, and dust waves. IV.]{Bow shocks, bow
  waves, and dust waves. IV. Bow shape statistics}

\newcommand\AddressCRyA{Instituto de Radioastronom\'{\i}a y Astrof\'{\i}sica,
  Universidad Nacional Aut\'onoma de M\'exico, Apartado Postal 3-72,
  58090 Morelia, Michoac\'an, M\'exico}
\author[Henney et al.]{
  William J. Henney,\thanks{w.henney@irya.unam.mx}
  Jorge A. Tarango-Yong,
  Luis \'Angel Guti\'errez-Soto,
  \& S. J. Arthur
  \\
  \AddressCRyA
}

% These dates will be filled out by the publisher
\date{Accepted XXX. Received YYY; in original form ZZZ}

% Enter the current year, for the copyright statements etc.
\pubyear{2019}

\input{bs-bw-dw-defs}

\begin{document}
\label{firstpage}
\pagerange{\pageref{firstpage}--\pageref{lastpage}}
\maketitle
\begin{abstract}
  Stellar bow shocks are the result of the supersonic interaction
  between a stellar wind and its environment.  Some of these are
  "runaways": high-velocity stars that have been ejected from a star
  cluster.  Others are "weather vanes", where it is the local
  interstellar medium itself that is moving, perhaps as the result of
  a champagne flow of ionized gas from a nearby \hii{} region.
  We recently proposed a new two-dimensional classification scheme for
  the shape of such bow shocks, which we here apply to three very
  different observational datasets: (1)~mid-infrared arcs
  around hot main-sequence stars (\(N = 227\)); (2)~far-infrared arcs
  around luminous cool stars (\(N = 16\)); and (3)~emission-line arcs
  around proplyds and other young stars in the Orion Nebula
  (\(N = 18\)).  We find significant differences between the three
  datasets: cool star bow shocks have markedly more closed wings than
  hot star bow shocks, while differences in the shape of the apex
  region are only marginally significant.  The Orion Nebula arcs, on
  the other hand, have both significantly more open wings and
  significantly flatter apexes than the hot star bow shocks.  We
  discuss the implications of these differences for understanding the
  physics of the bow shock interaction in the three classes of
  sources.
\end{abstract}

\begin{keywords}
  circumstellar matter -- methods: statistical -- stars: winds, outflows
\end{keywords}

\input{sec-obs-intro}
\input{sec-quadrics-observations}
\input{sec-standing-wave}
\input{sec-obs-conclusions}

\section*{Acknowledgements}
We are grateful for financial support provided by Dirección General de
Asuntos del Personal Académico, Universidad Nacional Autónoma de
México, through grants Programa de Apoyo a Proyectos de Investigación
e Inovación Tecnológica IN111215 and IN107019.

\bibliographystyle{mnras}
%% All references should be put in the BibTeX file bowshocks-biblio.bib
\bibliography{bowshocks-biblio}
\appendix
\input{app-p-values}

% Don't change these lines
\bsp	% typesetting comment
\label{lastpage}

### 2019-04-18 papers/Papers1234/sec-obs-intro.tex
Lots of reorganizing of Paper IV

@@

Instability of bow \citep{Blondin:1998a}

>>>>

In this paper, we investigate the shapes of stellar bow shocks by
calculating the distributions of planitude and alatude for different
classes of bow shock source.  The remainder of the paper is organized as follows: in \S~\ref{sec:mid-infrared-arcs} we present an analysis of the shapes of several hundred bowshock candidates associated with OB stars from the

@@

### 2019-04-18 papers/Papers1234/sec-quadrics-observations.tex
Lots of reorganizing of Paper IV

\section{Comparison with observations}
\label{sec:comp-with-observ}

Placing various classes of objects on the \(R_{90}\)--\(R_c\) plane:
\begin{itemize}
\item LL arcs
\item runaway O stars
\item AGB stars
\end{itemize}

\subsection{Mid-infrared arcs around early-type stars}
\label{sec:mid-infrared-arcs}

@@

\subsubsection{Automatic tracing and fitting of bow shocks}
\label{sec:autom-trac-fitt}

@@

\subsubsection{Subjective evaluation of the fit quality}
\label{sec:subj-eval-fit}

@@

\subsubsection{OB bow shock shapes on the diagnostic plane}
\label{sec:ob-shapes}

@@

\subsubsection{Correlation between bow shock size and other parameters}
\label{sec:corr-size}

@@

\subsubsection{Correlation between bow shock shape and other parameters}
\label{sec:corr-shape}
\begin{figure*}
  \centering
  \begin{tabular}{ll}
    (a) & (b) \\
    \includegraphics[width=0.45\textwidth]{figs/mipsgal-Rc-R90-R0} &
    \includegraphics[width=0.45\textwidth]{figs/mipsgal-Rc-R90-Mag} 
  \end{tabular}
  \caption[]{As Fig.~\ref{fig:mipsgal-uncorrelated} but for median
    splits of continuous parameters. (a)~Bow shock angular size,
    \(R_0\), divided into two equal-sized sub-samples: large (blue
    symbols and blue filled contours) and small (red symbols and red
    hollow contours).  (b)~Extinction-corrected \(H\)-band magnitude
    of the stellar source, divided into two equal-sized sub-samples:
    bright (red symbols and orange filled contours) and faint (green
    symbols and green hollow contours).}
  \label{fig:mipsgal-correlated}
\end{figure*}

@@

We now investigate if the bow shock shapes of the MIPSGAL sources are
correlated with any other parameters via the following methodology:
\begin{enumerate}[1.]
\item For each of the parameters in the \citet{Kobulnicky:2016a}
  catalog, we divide the sources into two or more sub-samples,
  according to the value of the parameter.  For quantitative
  parameters, such as those discussed in the previous section, we use
  two sub-samples of equal size, with membership determined by whether
  the parameter is larger or smaller than the median value.  But we
  also use categorical parameters, such as the type of bow shock
  environment, or the presence/absence of \SI{8}{\um} emission, in
  which case the sub-samples are of unequal size.
\item We plot the KDEs of the two sub-samples separately on the
  \(R_c\)--\(R_{90}\) plane (see Figs.~\ref{fig:mipsgal-correlated} and
  \ref{fig:mipsgal-uncorrelated}) to check for any obvious
  differences.
\item We check if there is any statistically significant difference
  between the bow shock shapes of the sub-samples by applying three
  different non-parametric tests to \(R_c/R_0\) and \(R_{90}/R_0\)
  separately.  The principal test used is the two-sample
  Anderson--Darling test \citep{Anderson:1952a, Scholz:1987a,
    Makarov:2017a}, which is a general test of the \textit{null
    hypothesis} that the two sub-samples are drawn from the same
  distribution.  It returns a \(p\)-value, which is the estimated
  probability that the observed difference between the two sub-samples
  would be as large as it is purely by chance if they were all were
  drawn from the same distribution.  We consider two different
  thresholds for significance: \(p < 0.05\) (approximately
  2-\(\sigma\) for a normal distribution) and \(p < 0.003\) (approximately
  3-\(\sigma\)).  We show in Appendix~\ref{sec:distr-p-values} that the
  more stringent \(p < 0.003\) condition is required in order to
  confidently reject the null hypothesis and declare a ``significant''
  difference between the two sub-samples, but we also consider the
  less strict threshold of \(p < 0.05\) as an indicator of
  ``possible'' difference.  We supplement the general-purpose
  Anderson--Darling test with two tests that probe specific features of
  the sample distributions: the Mann--Whitney--Wilcoxon \(U\) test
  \citep{Mann:1947a}, which is sensitive to differences in the central
  value (e.g., median) and the Brown--Forsythe test
  \citep{Brown:1974a}, which is sensitive to differences in the
  variance, or width, of the distribution (see
  Appendix~\ref{sec:distr-p-values} for details).
\end{enumerate}

@@

\subsection{Far-infrared arcs around late-type stars}
\label{sec:far-infrared-arcs}

@@

\subsection{Stationary emission line arcs in M42}
\label{sec:stat-emiss-line}

>>>>

\section{Mid-infrared arcs around early-type stars}
\label{sec:mid-infrared-arcs}
\label{sec:comp-with-observ}

@@

\subsection{Automatic tracing and fitting of bow shocks}
\label{sec:autom-trac-fitt}

@@

\subsection{Subjective evaluation of the fit quality}
\label{sec:subj-eval-fit}

@@

\subsection{OB bow shock shapes on the diagnostic plane}
\label{sec:ob-shapes}

@@

\subsection{Correlation between bow shock size and other parameters}
\label{sec:corr-size}

@@

\subsection{Correlation between bow shock shape and other parameters}
\label{sec:corr-shape}
\begin{figure*}
  \centering
  \begin{tabular}{ll}
    (a) & (b) \\
    \includegraphics[width=0.45\textwidth]{figs/mipsgal-Rc-R90-R0} &
    \includegraphics[width=0.45\textwidth]{figs/mipsgal-Rc-R90-Mag} 
  \end{tabular}
  \caption[]{As Fig.~\ref{fig:mipsgal-uncorrelated} but for median
    splits of continuous parameters. (a)~Bow shock angular size,
    \(R_0\), divided into two equal-sized sub-samples: large (blue
    symbols and blue filled contours) and small (red symbols and red
    hollow contours).  (b)~Extinction-corrected \(H\)-band magnitude
    of the stellar source, divided into two equal-sized sub-samples:
    bright (red symbols and orange filled contours) and faint (green
    symbols and green hollow contours).}
  \label{fig:mipsgal-correlated}
\end{figure*}

@@

We now investigate if the bow shock shapes of the MIPSGAL sources are
correlated with any other parameters via the following methodology:
\begin{enumerate}[1.]
\item For each of the parameters in the \citet{Kobulnicky:2016a}
  catalog, we divide the sources into two or more sub-samples,
  according to the value of the parameter.  For quantitative
  parameters, such as those discussed in the previous section, we use
  two sub-samples of equal size, with membership determined by whether
  the parameter is larger or smaller than the median value.  But we
  also use categorical parameters, such as the type of bow shock
  environment, or the presence/absence of \SI{8}{\um} emission, in
  which case the sub-samples are of unequal size.
\item We plot the KDEs of the two sub-samples separately on the
  \(R_c\)--\(R_{90}\) plane (see Figs.~\ref{fig:mipsgal-correlated} and
  \ref{fig:mipsgal-uncorrelated}) to check for any obvious
  differences.
\item We check if there is any statistically significant difference
  between the bow shock shapes of the sub-samples by applying three
  different non-parametric tests to \(R_c/R_0\) and \(R_{90}/R_0\)
  separately.  
\end{enumerate}

The principal test used is the two-sample Anderson--Darling test
\citep{Anderson:1952a, Scholz:1987a, Makarov:2017a}, which is a
general test of the \textit{null hypothesis} that the two sub-samples
are drawn from the same distribution.  It returns a \(p\)-value, which
is the estimated probability that the observed difference between the
two sub-samples would be as large as it is purely by chance if they
were all were drawn from the same distribution.  We consider two
different thresholds for significance: \(p < 0.05\) (approximately
2-\(\sigma\) for a normal distribution) and \(p < 0.003\) (approximately
3-\(\sigma\)).  We show in Appendix~\ref{sec:distr-p-values} that the more
stringent \(p < 0.003\) condition is required in order to confidently
reject the null hypothesis and declare a ``significant'' difference
between the two sub-samples, but we also consider the less strict
threshold of \(p < 0.05\) as an indicator of ``possible'' difference.
We supplement the general-purpose Anderson--Darling test with two tests
that probe specific features of the sample distributions: the
Mann--Whitney--Wilcoxon \(U\) test \citep{Mann:1947a}, which is
sensitive to differences in the central value (e.g., median) and the
Brown--Forsythe test \citep{Brown:1974a}, which is sensitive to
differences in the variance, or width, of the distribution (see
Appendix~\ref{sec:distr-p-values} for details).

@@

\section{Far-infrared arcs around late-type stars}
\label{sec:far-infrared-arcs}

@@

\section{Stationary emission line arcs in M42}
\label{sec:stat-emiss-line}
